---
---
title: "Final Project - SS"
output: html_document
author: "shannon spragg"
date: "10/22/21"
---
## Final Project

# Instructions:
Fit two different kinds of models (e.g. logistic regression and random forest; or, better yet, use the dismo vignette and use some of the additional classifiers there; or, better still, read some of the Introduction to Statistical Learning book and use classification approaches described there).

For each model-type, fit a global model (i.e., a model that contains all of the predictors you’ve developed) and a reduced model (i.e., a model with only the handful of predictors you imagine to be important based on your visualizations). You should have a total of 4 different model objects.

Use k-fold cross-validation (see the vignettes for dismo and Intro to Statistical Learning) to evaluate the preictive accuracy of each model


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Final Project: MT Wildlife Collisions
```{r load packages}
library(tidyverse)
library(pander)
library(sf)
library(terra)
library(raster)
library(dismo)
library(sp)
library(rgdal) 
library(rgeos)
library(pander)
library(randomForest)
library(rJava)
library(measurements)
install.packages("caret")
library(caret)
install.packages("ISLR2")
library(ISLR2)
library("boot")
```


```{r load data}
# Bring in Master DF and HMI raster:
mt.master.df <- st_read("data/mt_master_df.shp")
# add a column to show that the centroid points are presences
mt.master.df$y <- 1

human.mod.rast <- raster::raster('/opt/data/session08/hmi.tif')
mt.boundary <- tigris::states(cb=TRUE) %>% 
  filter(STUSPS == "MT") 
mt.pas <- st_read("data/MT_PAs.shp")

# Reproject Data:
mt.bound.reproj <- st_make_valid(mt.boundary) %>% 
  st_transform(crs=crs(human.mod.rast))
st_crs(mt.bound.reproj) # This worked, they match! Now, match master to this one
mt.master.reproj <- st_transform(mt.master.df, crs = st_crs(mt.bound.reproj))
st_crs(mt.master.reproj)
mt.pas.reproj <- st_transform(mt.pas, crs = st_crs(mt.bound.reproj))

#human.mod.crop <- terra::resample(human.mod.rast, mammal.tot.rich)

# Crop the HMI raster to MT:
hmi.mt.crop <- crop(human.mod.rast, mt.bound.reproj)
mt.outline <- st_boundary(mt.bound.reproj) # Just making this a boundary (not polygon)
hmi.mt.rast <- raster(hmi.mt.crop)
mt.pa.rast <- raster::rasterize(mt.pas.reproj, hmi.mt.crop)

pred.stack <- stack(hmi.mt.crop, mt.pa.rast)


# Plot these to check projections are correct:
plot(hmi.mt.crop)
plot(mt.outline, add = TRUE) 
plot(mt.master.reproj, add = TRUE) # This looks good!
```

```{r produce background pts}
# Now we produce pseudo-absence points:
abs <- sampleRandom(pred.stack, 4057, na.rm = TRUE, sp=TRUE)
plot(abs)
abs <- st_as_sf(abs)
abs$y <- 0
colnames(abs)

# give it an ID column:
abs <- dplyr::mutate(abs, ID = row_number())
abs <- abs %>% 
  relocate(ID)
```

```{r add predictors to abs pts}
# Create a PA variable:
abs.pas <- abs %>%
  st_join(., mt.pas.reproj[, c("d_Mang_Typ", "d_Mang_Nam", "Unit_Nm" , "Loc_Nm")], left = TRUE) 

abs.pas.merged <- abs.pas %>%
  distinct(ID, .keep_all = TRUE) # Here we just drop the duplicates

# Let's create a Dist to PA's Column and add to the df ------------------------
# Calculation of the distance between the PA's and our points:
dist.col2pas <- st_distance(abs.pas.merged, mt.pas.reproj)
head(dist.col2pas)
# Must find the minimum distance to PA's (Distance from conflict point to nearest PA)
min.dist <- apply(dist.col2pas, 1, min)
str(min.dist)
# Add Distance Variable into Datatable:
abs.pas.merged$distance_to_PAs<-min.dist
# Make the Dist to PA's in km:
dist_in_km<-conv_unit(abs.pas.merged$distance_to_PAs,"m","km")
str(dist_in_km)
abs.pas.merged$distance_to_PAs<-dist_in_km
head(abs.pas.merged) # Looks good!
```


```{r add predictors to pres pts}
# Extract the pres stack to pres points:
mt.master.sp <- as(mt.master.reproj, 'Spatial') # Making master.df sp object

pres.pts <- extract(pred.stack, mt.master.sp, df=TRUE, na.rm = TRUE)

# Bind our presence points /extracted values to the master df:
pres.pts.df <- cbind(mt.master.reproj, pres.pts)

# Add variables to our pres points:
pres.pts.subset <- pres.pts.df %>%
  dplyr::select(ID, hmi, layer, y)

# Create a PA variable:
pres.pas <- pres.pts.subset %>%
  st_join(., mt.pas.reproj[, c("d_Mang_Typ", "d_Mang_Nam", "Unit_Nm" , "Loc_Nm")], left = TRUE) 

pres.pas.merged <- pres.pas %>%
  distinct(ID, .keep_all = TRUE) # Here we just drop the duplicates

# Let's create a Dist to PA's Column and add to the df ------------------------
# Calculation of the distance between the PA's and our points:
pdist.col2pas <- st_distance(pres.pas.merged, mt.pas.reproj)
head(pdist.col2pas)

# Must find the minimum distance to PA's (Distance from conflict point to nearest PA)
pmin.dist <- apply(pdist.col2pas, 1, min)
str(pmin.dist)

# Add Distance Variable into Datatable:
pres.pas.merged$distance_to_PAs<-pmin.dist

# Make the Dist to PA's in km:
pdist_in_km<-conv_unit(pres.pas.merged$distance_to_PAs,"m","km")
str(pdist_in_km)
pres.pas.merged$distance_to_PAs<-pdist_in_km
head(pres.pas.merged) # Looks good!

```

```{r join pres and abs}
# Now we join the pres and abs dataframes:
pres.abs.df <- rbind(abs.pas.merged, pres.pas.merged)
head(pres.abs.df)
# Now these are in single df with pres and abs and our predictors
```


```{r scale predictors}
# Make sure our predictors are near each other:
pres.abs.df <- pres.abs.df %>% relocate(distance_to_PAs, .before = y)

#Now let's scale the predictors: (have to scale individually bc won't work all together)
pres.abs.df$hmi <- scale(pres.abs.df$hmi)
pres.abs.df$layer <- scale(pres.abs.df$layer)
pres.abs.df$distance_to_PAs <- scale(pres.abs.df$distance_to_PAs)

```

```{r model one - logistic regression}
logistic.global <- glm(y ~ hmi + distance_to_PAs + layer, family=binomial(link="logit"), data = pres.abs.df)
logistic.reg1 <- glm(y ~ hmi + layer, family = binomial(link="logit"), data = pres.abs.df)
```

```{r model two - random forest}
reg.model <- y ~ hmi + distance_to_PAs + layer
rf1 <- randomForest(reg.model, data=pres.abs.df, na.action=na.exclude)
## Warning in randomForest.default(m, y, ...): The response has five or fewer
## unique values. Are you sure you want to do regression?
class.model <- factor(y) ~ hmi + layer + distance_to_PAs
rf2 <- randomForest(class.model, data=pres.abs.df, na.action=na.exclude)
# Look at predictor importance based on splits in decision tree that uses the variables:
varImpPlot(rf1)
varImpPlot(rf2)

# It looks like variable hmi and layer are more pure than the other variable, which is pretty close to zero
```

```{r k fold cross validation}
# Let's try this k - fold cross validation:
pres.abs.no.geo <- st_drop_geometry(pres.abs.df)

set.seed (1)
# Split observations into training set:
train <- sample (4057, 2028)
lm.fit <- lm(y~ hmi, data = pres.abs.df , subset = train)
# Estimate response for observations:
attach(pres.abs.df)
mean((y - predict(lm.fit , pres.abs.df))[-train ]^2)
# The estimated test MSE for the linear regression fit is 0.6666119

# Use the poly() function to estimate the test error for the quadratic and cubic regressions:
lm.fit2 <- lm(y~ poly(hmi, 2), data = pres.abs.df , subset = train)

mean((y - predict(lm.fit2, pres.abs.df))[-train]^2)
# [1] 18.72
lm.fit3 <- lm(y~ poly(hmi, 3), data = pres.abs.df , subset = train)
mean((mpg - predict (lm.fit3 , pres.abs.df))[-train]^2)

# Leave one out cross validation:
glm.fit <- glm (y ~ hmi , data = pres.abs.df)
coef(glm.fit)

glm.fit <- glm(y ~ hmi , data = pres.abs.df)
cv.err <- cv.glm(pres.abs.df , glm.fit)
cv.err$delta
# [1] 0.1259023 0.1259023 These are our cross validation results with a K of 8114


```
